---
title: "r2dii.data 0.1.2 and r2dii.match 0.0.4 are now on CRAN"
description: |
  Less bugs and more features
author:
  - name: Mauro Lepore
    url: https://github.com/maurolepore
base_url: https://2degreesinvesting.github.io/
date: 2020-08-14
output: 
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: true
categories:
  - r2dii
  - software development
  - release
preview: preview.jpg
twitter:
  site: "@mauro_lepore"
  creator: "@mauro_lepore"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

r2dii.data 0.1.2 and r2dii.match 0.0.4 are now on CRAN. These packages provide datasets and tools to align financial markets to climate goals. These releases fix a number of bugs that you can learn about [here](https://2degreesinvesting.github.io/r2dii.data/news/index.html) and [here](https://2degreesinvesting.github.io/r2dii.match/news/index.html); this post shows enhancements and new features.

You can install r2dii.data and r2dii.match from CRAN with:

```r
install.packages("r2dii.data")
install.packages("r2dii.match")
```

Then use them with:

```{r}
library(r2dii.data)
library(r2dii.match)
```

### r2dii.data 0.1.2

r2dii.data 0.1.2 includes two new dataset -- [`green_or_brown`](https://2degreesinvesting.github.io/r2dii.data/reference/green_or_brown.html), and [`sic_classification`](https://2degreesinvesting.github.io/r2dii.data/reference/sic_classification.html) (thanks to [Daisy Pacheco](https://github.com/daisy-pacheco) and [George Harris](https://github.com/georgeharris2deg)).

```{r}
green_or_brown

sic_classification
```

Also, [`region_isos`](https://2degreesinvesting.github.io/r2dii.data/reference/region_isos.html) gained data from ETP 2017, and [`ald_demo`](https://2degreesinvesting.github.io/r2dii.data/reference/ald_demo.html) dropped the column `number_of_assets` (thanks to [Taylor Posey](https://github.com/tposey28)).

```{r}
unique(region_isos$source)

any(grepl("number_of_assets", names(ald_demo)))
```

### r2dii.match 0.0.4

`match_name()` now outputs a new column -- `borderline`. This column helps you measure how much of your loanbook matched some asset; see the new article [Calculating matching coverage](https://2degreesinvesting.github.io/r2dii.match/articles/matching-coverage.html).

```{r}
loanbook <- loanbook_demo
ald <- ald_demo

matched <- match_name(loanbook, ald)
tail(names(matched))
```

Also, `match_name()` now runs faster and uses less memory. This responds to users's feedback, diligently managed by [George Harris](https://github.com/georgeharris2deg) -- thanks! If you still run out of memory, read [Using `match_name()` with large loanbooks: How to resolve memory issues](https://2degreesinvesting.github.io/posts/2020-07-31-chunk-your-data/) and [Improving r2dii.match: How to work with big data, and benchmarks of a more efficient version of `match_name()`](https://2degreesinvesting.github.io/posts/2020-07-18-improving-r2dii-match/). You may also want to reduce the size of your data: use the new function [`crucial_lbk()`](https://2degreesinvesting.github.io/r2dii.match/reference/crucial_lbk.html) to select the minimum columns you need for `match_name()`.

```{r}
ncol(loanbook)

crucial_lbk()

smaller_loanbook <- loanbook[crucial_lbk()]
ncol(smaller_loanbook)

match_name(smaller_loanbook, ald)
```
## Acknowledgements

```{r Acknowledgements, echo=FALSE}
# bash
# git shortlog --summary tag-oldrel..tag-release
# For example: Who contributed to r2dii.match 0.0.2 (since 0.0.1)?
# git shortlog --summary 0.0.1..0.0.2

thank <- function(x) {
  x <- sort(x)
  paste0("[", x, "]", sprintf("(https://github.com/%s)", x), collapse = ", ")
}
contributors <- c(
  "jdhoffa",
  "maurolepore"
)
```

While this release includes commits from only a few of us (`r thank(contributors)`), it is thanks to feedback from our colleagues and users.